Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-04-26 files/Chapter2/generalvibrationmodel.tex
Chapter 3

\subsection{General vibration problem new}
\textcolor{red}{Introduction***}

Consider Problem GVar defined in Section \ref{ssec:existence:VariationlApproach}. In this section, assume there is no damping and no forcing.

@@

A solution to the eigvalue problem consists of an eigenvalue $\lambda$ with corresponding eigenvector $x$.

@@

Since this differential equation is a simple second order differential equation, T(t) has the following possible solutions:
\begin{flalign*}
	T_n(t) &= e^{-r_nt}\left[ A_n \cos(\omega_n t) + B_n \sin(\omega_n t) \right],\\
	T_n(t) &= e^{-r_nt}\left[ A_n + B_n t) \right],\\
	T_n(t) &= e^{-r_nt}\left[ A_n \cosh(\mu_n t) + B_n \sinh(\mu_n t) \right].
\end{flalign*}

Then the formal series solution for the boundary value problem is
\begin{eqnarray}
	u(t) = \sum_{n=1}^{\infty} T_n(t)x_n. \label{eq:1D_Model:ModalAnalysisSeriesSolution}
\end{eqnarray}

>>>>

\subsection{General vibration problem new}
This section is a discussion of a general vibration problem. The discussion will refer to the article \cite{CVV18}. Consider Problem GVar defined in Section \ref{ssec:existence:VariationlApproach}. The authors of \cite{CVV18} do take damping into consideration, which is not covered in this dissertation. Assume there is no damping and no forcing.

@@

A solution to the eigenvalue problem consists of an eigenvalue $\lambda$ with corresponding eigenvector $x$.

@@

Since this differential equation is a simple second order differential equation, T(t) has the following possible solutions as given in \cite{CVV18}:
\begin{flalign}
	T_n(t) &= e^{-r_nt}\left[ A_n \cos(\omega_n t) + B_n \sin(\omega_n t) \right] \quad \textrm{ if } \lambda > 0, \label{lambda_1}\\
	T_n(t) &= e^{-r_nt}\left[ A_n + B_n t \right] \quad \textrm{ if } \lambda = 0, \label{lambda_2}\\
	T_n(t) &= e^{-r_nt}\left[ A_n \cosh(\mu_n t) + B_n \sinh(\mu_n t) \right] \quad \textrm{ if } \lambda < 0. \label{lambda_3}
\end{flalign}

From assumptions \textbf{A2} and \textbf{A3}, the bilinear forms $b$ and $c$ are positive definite. And since we assumed that there is no damping ($a=0$), $\lambda$ cannot be negative. Therefore, the only possible solutions for $T_n(t)$ is \eqref{lambda_1} and \eqref{lambda_2}.

Then combining the solutions of the eigenvalue problem and the differential equation, the formal series solution for the boundary value problem is
\begin{eqnarray}
	u(t) = \sum_{n=1}^{\infty} T_n(t)x_n. \label{eq:1D_Model:ModalAnalysisSeriesSolution}
\end{eqnarray}

### 2023-04-26 files/Chapter3/dynamicconvergence.tex
Chapter 3

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, two convergence results for the Finite Element Method (FEM) is discussed. The first result is the convergence of the Galerkin approximation for second order hyperbolic type problems, from the article \cite{BV13}. The second result is from the textbook \cite{SF97} regarding the convergence of the eigenvalues and eigenfunctions for a one-dimensional vibration problem, using the Finite Element Method.

\section{Convergence of the Galerkin approximation}
In the article \cite{BV13}, the authors investigate the convergence of the Galerkin approximation for second order hyperbolic type problems. The article \cite{BV18} extends \cite{BV13} by including general damping and damping at the endpoints. For the models in Chapter 1, \cite{BV13} is sufficient and \cite{BV18} was read for more insight and improved notation.\\

In Section 2.2, a general variational problem, called Problem GVar, which is identical to the one in \cite{VV02} is given. From this model, a Galerkin Approximation can be derived.

\subsubsection*{Problem GVar}
Given a function $f:J\rightarrow X$, find a function $u\in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$ and for each $t\in J,\ u(t)\in V,\ u'(t) \in V,\ u''(t)\in W$ and
\begin{eqnarray}
	c(u''(t),v)+a(u'(t),v)+b(u(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in V, \label{GV_PB1}
\end{eqnarray}
while $u(0)=u_{0}$, $u'(0)=u_{1}$.\\

It is assumed that the assumptions \textbf{A1}-\textbf{A4} from Section 2.2 are satisfied such that Problem GVar has a unique solution.\\

Before the main theory of \cite{BV13} can be discussed, some preparation is required. First the Galerkin approximation is for Problem GVar is derived. Then a equivalent system of ordinary differential equations is derived using the Finite Element Method.\\

\subsection{Galerkin approximation}
Recall from Chapter 2 that $V$ is the closure of the test function space $T$ in the Hilbert space $H$.\\

Suppose $S^h$ is a finite n-dimensional subspace of V. Let $\{\phi_1, \ \phi_2,...,\phi_n\}$ be a set of linear independent basis functions such that $S^h = \textrm{span}(\{\phi_1, \phi_2, ..., \phi_n \})$ and $\phi_i \in C^{(1)}$ for all $i$.\\

@@

\subsubsection*{Problem GVar$^h$}
Given a function $f: J \rightarrow X$, find a function $u_h \in C^2(J)$ such that for each $t\in J$
\begin{eqnarray}
	c(u_h''(t),v)+a(u_h'(t),v)+b(u_h(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in S^h, \label{DC_E2}
\end{eqnarray}
with the initial values $u_h(0)=u^h_{0}$ and $u'_h(0)=u^h_{1}$. The initial conditions $u^h_{0}$ and $u^h_{1}$ are projections of $u_0$ and $u_1$ into the finite dimensional space $S^h$.\\

@@

This theorem is presented in \cite{BV13}.

@@

Using these matrices, Problem GVar$^h$ is rewritten as a system of ordinary differential equations denoted by Problem ODE.
\subsubsection*{Problem ODE}
Find a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\bar{u}'' + C \bar{u}' + K\bar{u} = F(t) \ \ \ \textrm{ with } \bar{u}(0) = \bar{u}^h_0 \textrm{ and }  \bar{u}(1) = \bar{u}^h_1 \label{ODE}
\end{eqnarray}
The function $F$ is the interpolation of the function $f$ into the discrete space $S^h$.\\

The following propositions for Problem ODE are given in \cite{BV13}.

@@

Since the set $\{\phi_1, \ \phi_2,...,\phi_n\}$ is a linearly independent basis for $S^h$, the bilinear form $c$ is positive definite on $S^h$. Therefore, the matrix $M$ is invertible. The equation \eqref{ODE} of Problem ODE can be rewritten as $\bar{u}'' + M^{-1}C\bar{u}' + M^{-1}K \bar{u} = M^{-1}F(t)$. Using the theory of linear differential equations, a unique solution for Problem ODE can be found. This proves Proposition 1.\\

@@

\subsection{Error estimates}\label{e_est}
Given that the assumptions \textbf{A1}-\textbf{A4} of Section 2 is satisfied, and $f$ is continuous, there exists unique solutions for  Problem GVar and Problem GVar$^h$. It remains only to show that the solution of Problem GVar$^h$ denoted by $u^h$ converges to the solution of Problem GVar denoted by $u$ as $h$ is decreased.\\

Suppose $u$ is a solution of Problem GVar and $u^h$ is a solution of Problem GVar$^h$. Define the error $e^h$ by
\begin{eqnarray*}
	e^h(t) = u(t) - u^h(t).
\end{eqnarray*}

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, two important convegence results of the Finite Element Method (FEM) is discussed. The first result looks at the convergence of the Galerkin approximation for second order hyperbolic type problems, as presented in \cite{BV13}. The second result, from the textbook \cite{SF97}, examines the convergence of eigenvalues and eigenfunctions for a one-dimensional vibration problem, using the Finite Element Method.

\section{Convergence of the Galerkin approximation}
In the article \cite{BV13}, the authors investigate the convergence of the Galerkin approximation for second order hyperbolic type problems. The article \cite{BV18} extends the work of \cite{BV13} by including general damping and damping at the endpoints. For the models in Chapter 1, \cite{BV13} is sufficient while \cite{BV18} provides more insight and improved notation.

In Section 2.2, Problem GVar is presented and identical to the problem in \cite{BV13}. The problem is repeated here for convenience.

\subsubsection*{Problem GVar}
Given a function $f:J\rightarrow X$, find a function $u\in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$ and for each $t\in J,\ u(t)\in V,\ u'(t) \in V,\ u''(t)\in W$ and
\begin{eqnarray}
	c(u''(t),v)+a(u'(t),v)+b(u(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in V, \label{GV_PB1}
\end{eqnarray}
while $u(0)=u_{0}$, $u'(0)=u_{1}$.

Assume that the assumptions \textbf{A1}-\textbf{A4} from Section 2.2 are satisfied ensuring that Problem GVar has a unique solution.

Before the theory of \cite{BV13} can be discussed, some preliminary work is necessary. The structure of this section is as follows. First the Galerkin approximation is for Problem GVar is derived. Then a equivalent system of ordinary differential equations is derived using the Finite Element Method. Finally, the convergence of the Galerkin approximation is discussed using the work of the article \cite{BV13}.

\subsection{Galerkin approximation of Problem GVar}
Recall from Chapter 2 that $V$ is the closure of the test function space $T$ in the Hilbert space $H$.

Suppose $S^h$ is a finite n-dimensional subspace of V. Let $\{\phi_1, \ \phi_2,...,\phi_n\}$ be a set of linear independent basis functions such that $S^h = \textrm{span}(\{\phi_1, \phi_2, ..., \phi_n \})$ and $\phi_i \in C^{(1)}$ for all $i$.

@@

\subsubsection*{Problem GVar$^h$}
Given a function $f: J \rightarrow X$, find a function $u_h \in C^2(J)$ such that for each $t\in J$
\begin{eqnarray}
	c(u_h''(t),v)+a(u_h'(t),v)+b(u_h(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in S^h, \label{DC_E2}
\end{eqnarray}
with the initial values $u_h(0)=u^h_{0}$ and $u'_h(0)=u^h_{1}$. The initial conditions $u^h_{0}$ and $u^h_{1}$ are projections of $u_0$ and $u_1$ into the finite dimensional space $S^h$.

@@

This theorem is presented and proved in \cite{BV13}.

@@

Using these matrices, Problem GVar$^h$ is rewritten as a system of ordinary differential equations denoted by Problem ODE.
\subsubsection*{Problem ODE}
Find a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\bar{u}'' + C \bar{u}' + K\bar{u} = F(t) \ \ \ \textrm{ with } \bar{u}(0) = \bar{u}^h_0 \textrm{ and }  \bar{u}(1) = \bar{u}^h_1 \label{ODE}
\end{eqnarray}
The function $F$ is the interpolation of the function $f$ into the discrete space $S^h$.

The following propositions related to Problem ODE are given in \cite{BV13}.

@@

Since the set $\{\phi_1, \ \phi_2,...,\phi_n\}$ forms a linearly independent basis for $S^h$, the bilinear form $c$ is positive definite on $S^h$. Consequently, the matrix $M$ is invertible. The equation \eqref{ODE} can therefore be rewritten as $\bar{u}'' + M^{-1}C\bar{u}' + M^{-1}K \bar{u} = M^{-1}F(t)$. Using the theory of linear differential equations, a unique solution for Problem ODE can be found, proving Proposition 1.

@@

\subsection{Error estimates}\label{e_est}
Given that the assumptions \textbf{A1}-\textbf{A4} of Section 2 is satisfied, and $f$ is continuous, there exists unique solutions for  Problem GVar and Problem GVar$^h$. It remains only to show that the solution of Problem GVar$^h$ denoted by $u^h$ converges to the solution of Problem GVar denoted by $u$.

Let $u$ be the solution of Problem GVar and $u^h$ be the solution of Problem GVar$^h$. The authors of \cite{BV13} define the following error,

\begin{eqnarray}
	e^h(t) = u(t) - u^h(t). \label{Error}
\end{eqnarray}

### 2023-04-26 files/Chapter4/ModalAnalysisTimo.tex
Chapter 3

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
An example of modal analysis on a Timoshenko beam is given in section \ref{sec:existence:ModalAnalysis}. In this section, more general theory for modal analysis of te Timoshenko beam theory is discuss, as well as some more examples of the application of the theory.
\section{Eigenvalue problem}\label{sec:Timo:EigenvalueProblem}
This section is a discussion of a systematic method to solve the eigenvalue problem for the Timoshenko beam theory. The article discussed is \cite{VV06}.

@@

@@

From the general solutions for $u$ and $\phi$, the eigenvalues and eigenfunctions can be determined by imposing the boundary conditions of the model problem. The boundary conditions at $x=0$ reduces the solution space of four dimensions to a solution space of dimension at least two. The boundary condition at $x=1$ is then substituted which results in a homogeneous system of linear equations.
\begin{eqnarray*}
A \bar{b} = \bar{0}
\end{eqnarray*}

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
An example of modal analysis on a Timoshenko beam is given in section \ref{sec:existence:ModalAnalysis}. In this chapter, more general theory for modal analysis of te Timoshenko beam theory is discussed, as well as some more examples of the application of the theory.

\section{Eigenvalue problem}\label{sec:Timo:EigenvalueProblem}
This section is a discussion of a systematic method to solve the eigenvalue problem for the Timoshenko beam theory. The article discussed is \cite{VV06}.

@@

\textcolor{red}{*****************SUGGESTION****************************}

\subsubsection*{Stragety for determining the eigenvalues and eigenvectors}

The authors of \cite{VV06} provide a detailed example of the application of the above strategy to for determining the eigenvalues and eigenvectors. The example used is a pinned-pinned beam, the same model as in section \ref{subsec:Timoshenko_Modal_Analsis}.

The strategy can be summarized as follows:

From the general solutions for $u$ and $\phi$, the eigenvalues and eigenfunctions can be determined by imposing the boundary conditions at $x=0$ to reduce the solution space of four dimensions to a solution space of dimension at least two.

The boundary condition at $x=1$ is then substituted which results in a homogeneous system of linear equations of the form
\begin{eqnarray*}
A \bar{b} = \bar{0}.
\end{eqnarray*}

This system either has the zero solution or infinitely many solutions. To ensure the zero solution, the determinant of the matrix $A$ is set to zero. This equation $\textrm{det}(A) = 0$ is called the frequency equation.

The frequency equation has infinitely many solutions. Each of the solutions corresponds to a eigenvalue with a unique vector $\bar{b}$. The eigefunctions can then be obtained by substituting the vector $\bar{b}$ into the general solution for $u$ and $\phi$.

\subsubsection*{Pinned-pinned beam}

Returning back to the example of the pinned-pinned beam, we present some of the results from \cite{VV06}:

For $\lambda < \alpha$ the general solution reduces to $\langle u(x) , \phi(x) \rangle = \langle \sin(k \pi x), A_k \cos(k \pi x) \rangle$.

For $\lambda = \alpha$, the general solution reduces to $\langle u(x) , \phi(x) \rangle = \langle 0, 1 \rangle$.

For $\lambda > \alpha$, the general solution also reduces to $\langle u(x) , \phi(x) \rangle = \langle \sin(k \pi x), A_k  \cos(k \pi x) \rangle$.

A rigourous proof of the above results can be found in \cite{VV06}. These results also prove the Theorem 1 presented in section \ref{subsec:Timoshenko_Modal_Analsis}.

\textcolor{red}{*****************OLD alles onder moontlik weg vat****************************}

@@

From the general solutions for $u$ and $\phi$, the eigenvalues and eigenfunctions can be determined by imposing the boundary conditions at $x=0$ to reduce the solution space of four dimensions to a solution space of dimension at least two.

The boundary conditions at $x=1$ is then substituted which results in a homogeneous system of linear equations of the form
\begin{eqnarray*}
A \bar{b} = \bar{0}
\end{eqnarray*}
This system either has the zero
